From Nested Virtualization Architecture \\To 2D Security Management for Distributed Clouds

### 2014-03-23 paper/draft.tex
Updating nv section

Nested virtualization refers to a system architecture with two layers of virtualization. It consists in a ``virtualized'' virtualization, where the guest system virtualize a nested guest. The concept could be generalized to an arbitrary number of nested guest layers, introducing recursive virtualization. Virtualization is able to execute tasks in a deprivileged context that normally requires a privileged execution. With nested virtualization we could extend this approach to virtualization itself, executing a guest hypervisor with nested guest VMs, without requiring any change to any of them.
%More formally, Popek \& Golberg \cite{Popek:1974:FRV:361011.361073} discussed this subject, introducing the concept of ``recursively virtualizable'' machine and giving an initial definition of recursive virtual machines: If it is possible for a virtual machine to run under itself a copy of the VMM that could also exhibit all the properties of the VMM and this procedure could be repeated until the resources of the systems are consumed, then the original machine is recursively virtualizable.

Introducing nested virtualization, we have to identify the two main actors involved in the virtualization process: bare-metal hypervisor (L0) and nested hypervisor (L1). In addition, we will refer to the nested guest as L2 guest. 
However, virtualization introduces a computational overhead. It is normally acceptable when we are dealing with single layer virtualization, but it grows exponentially when we try to add further levels \cite{rec:virt}. This huge performance overhead has prevented nested virtualization adoption during last years and had been finally addressed by Ben Yehuda et Al \cite{turtle:ibm} proposing an architecture with acceptable overhead.

Nested virtualization feasibility opened a plenty of new possibilities and perspectives, that have been largely discussed in literature \cite{rec:virt,turtle:ibm,art:blan}. The first aspect in common among different works proposed in literature is to diversify hypervisor functions, introducing new features but keeping the existing ones. In a general analysis of existing works, it could be possible to identify the general role of each different layer.
On one hand, L0 is charged of the vertical aspects of the platform. It has to achieve  the best possible performances and enforcing the strongest possible level of isolation. L0 represents the last line of defense of the platform and a privileged point of global status observation.

On the other hand, L1 is responsible for horizontal feature set. It has to provide the widest possible support to different platforms and the capacity to virtualize adopting different techniques (para-virtualization, hardware-assisted, dynamic binary translation). According to these requirements, Williams et al. \cite{art:blan} propose a patched version of Xen, Xen Blanket, to accomplish a L1 virtualization layer, designed to run over different platforms. This concretely allows to run over different physical platform the same VM, and live-migrate it at the occurrence.

>>>>

\subsection{Principle}

\noindent \textit{Nested virtualization (NV)} refers to a system architecture with two layers of virtualization~\cite{turtle:ibm}. It consists in ``virtualized'' virtualization: the guest system virtualizes a nested guest. The concept may be generalized to an arbitrary number of nested guest layers, leading to recursive virtualization~\cite{rec:virt,Popek:1974:FRV:361011.361073}. (Full) virtualization is able to execute tasks in a deprivileged context that normally requires a privileged execution. With NV, this approach may be extended to virtualization itself: a guest hypervisor may be run with nested guest VMs without requiring any change to any of them.
%More formally, Popek \& Golberg \cite{Popek:1974:FRV:361011.361073} discussed this subject, introducing the concept of ``recursively virtualizable'' machine and giving an initial definition of recursive virtual machines: If it is possible for a virtual machine to run under itself a copy of the VMM that could also exhibit all the properties of the VMM and this procedure could be repeated until the resources of the systems are consumed, then the original machine is recursively virtualizable.

Two layers should be distinguished when migrating to an NV architecture: (1) the hypervisor running above the hardware (\textit{L0 hypervisor}); and (2) the nested hypervisor (\textit{L1 hypervisor}). Nested guests are generally referred to as \textit{L2 guests}.
Unfortunately, virtualization comes with a price due to the extra layer of indirection introduced: the computational overhead is generally acceptable when dealing with single-layer virtualization, but grows exponentially when adding further levels~\cite{rec:virt}. This huge performance overhead prevented NV adoption during many years. However, Ben Yehuda et al. finally proposed an architecture tending towards more reasonable overheads~\cite{turtle:ibm}.

NV feasibility opened plenty of new possibilities and avenues for research that have been largely discussed in the literature~\cite{turtle:ibm,rec:virt,art:blan}. Such works all tend to diversify hypervisor functionalities, introducing new features but also keeping existing ones. It turns out that each layer addresses clearly different sets of issues.

L0 tackles vertical aspects of the platform. It aims to guarantee the best performance and strongest level of isolation possible. L0 represents the last line of defense of the platform and a privileged point for monitoring of its global status.

L1 is responsible for horizontal feature set. It has to provide the widest possible support for different platforms and the ability to virtualize using different techniques e.g., para-virtualization, hardware-assisted virtualization, dynamic binary translation. {\sffamily To address such requirements, Williams et al.~\cite{art:blan} proposed a patched version of Xen, Xen Blanket, to realize an L1 virtualization layer designed to run over different hypervisors. This notably enables to run and to perform live migration of the same VM over different physical cloud infrastructures.} % A migrer dans le related work?
